---
title: "Introduction to Machine Learning - FIFA Case"
author: "A. Piguet - L. Gatinet - J. Guillot"
output: html_notebook
---

```{r, echo=FALSE, eval=TRUE}
getwd()
setwd("/Users/antoinepiguet/Desktop/EDHEC DAAI/Introduction to Machine Learning")
```

```{r, echo=FALSE, eval=TRUE}
library(Amelia)
library(readr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(randomForest)
library(shiny)
```

```{r, eval=TRUE}
data <- read_csv("fifadata.csv") #is it possible to use a link here to make it really easy to read in the end ?

data_test <- read_csv("fifadata.csv")
```
***
# I-Data cleaning using dplyr:

```{r, eval=TRUE}
data <- data %>% select(-Photo, -Flag, -`Club Logo`, -(LS:RB), -Joined, -`Loaned From`, -X1, -Special, -`Real Face`, -`Body Type`, -`Jersey Number`)
```
Removing the Photo, Flag, Club Logo, Loaned From columns // We removed scores of each players for each position on the field
Body Type removed because of high correlation with Height and Weight


#### Converting arrival date in R format
```{r, eval=TRUE}
#data[which(is.na(data[,"Joined"])),"Joined"] <- "Loaned" # C'est tout casse
#data$Joined <- strptime(data$Joined, format="%b %d, %Y")
```

Replacing the NA's in the Joined column by the string "Loaned"


#### Converting financial values as integers (Wage, Release Clause and Value)
```{r}
data$Value <- ifelse(substring(data$Value, nchar(data$Value)) == "M",
                     as.numeric(substring(data$Value,2,nchar(data$Value)-1)),
                     as.numeric(substring(data$Value,2,nchar(data$Value)-1)) / 1000)
data$Wage <- as.numeric(substring(data$Wage,2,nchar(data$Wage)-1))
data$`Release Clause` <- as.numeric(substring(data$`Release Clause`,2,nchar(data$`Release Clause`)-1))
```

#### Converting measures of height and weight into international metric system
```{r}
data$Weight <- as.numeric(substring(data$Weight, 1, 3)) * 0.45359237
data <- data %>% separate(Height, c('feet', 'inches'), "'", convert = TRUE) %>%
  mutate(Height = (12*feet + inches)*2.54) %>% select(-inches, -feet)
```

#### Renaming all modified variables
```{r}
data <- data %>% rename(`Value (in M€)` = Value,
                        `Wage (in K€)` = Wage,
                        `Release Clause (in M€)` = `Release Clause`,
                        `Weight (in kg)` = Weight,
                        `Height (in cm)` = Height)
```

#### Splitting the Work Rate column into its 2 components (Offensive / Defensive)
```{r}
data <- data %>% separate(`Work Rate`, c('Offensive Work Rate', 'Defensive Work Rate'), "/ ")
data$`Offensive Work Rate` <- factor(data$`Offensive Work Rate`, levels=c("Low", "Medium", 'High'))
data$`Defensive Work Rate` <- factor(data$`Defensive Work Rate`, levels=c("Low", "Medium", 'High'))
data$Field_Position <- factor(data$Field_Position, levels=c("Goalkeeper", "Defenser", 'Mitfielder', "Attack"))
data$`Preferred Foot` <- factor(data$`Preferred Foot`, levels=c("Left", "Right"))
data$`International Reputation` <- factor(data$`International Reputation`, levels=1:5)
data$`Weak Foot` <- factor(data$`International Reputation`, levels=1:5)
data$`Skill Moves` <- factor(data$`International Reputation`, levels=1:5)
data$Club <- as.factor(data$Club)

summary(data$`International Reputation`)
```

#### Creating a column that captures the remaining duration of the player's contract

```{r}
text <- NULL
first <- NULL
last <- NULL

for (i in 1:nrow(data)) {
  text[i]=data$`Contract Valid Until`[i] 
  first[i]=nchar(text[i])-4
  last[i]=nchar(text[i])
  data$`Contract Valid Until`[i]=substring(text[i],first[i],last[i])
}

data$`Remaining Contract Duration` <- as.numeric(data$`Contract Valid Until`) - 2018

data <- data %>% select(-`Contract Valid Until`)
```


```{r}
for (i in 1:nrow(data)) {
  if (data$Position[i] %in% c("LW", "LF", "RW", "RF", 'CF', 'LS', 'RS', 'ST')) data$Field_Position[i] <- "Attack"
  if (data$Position[i] %in% c('CDM', 'LDM', 'RDM', 'CM', 'LCM', 'RCM', 'CAM', 'LAM', 'RAM', 'LM', 'RM')) data$Field_Position[i] <- "Mitfielder"
  if (data$Position[i] %in% c('CB', 'LCB', 'RCB', 'LB', 'RB', 'LWB', 'RWB')) data$Field_Position[i] <- "Defenser"
  if (data$Position[i] %in% c("GK")) data$Field_Position[i] <- "Goalkeeper"
}

data <- data %>% select(-Position)
```

```{r}
df <- data[which(apply(is.na(data),1,any)),]
```

All remaining NA's are players who have been loaned to another team (1264 players) 

nad 300 other players for whom FIFA did not gather the information or was denied the access

```{r}
data <- na.omit(data)
```
***
# II-Data Exploration using dplyr & ggplot

```{r}
ggplot(data) + aes(x=`Wage (in K€)`) + geom_histogram(bins=30) + scale_x_log10()
```

```{r}
datareg <- data %>% select(-ID, -Name, -Nationality, -Club) # Removing irrelevant categorical variables for the regression
my.model <- lm(`Value (in M€)`~., data=datareg)
summary(my.model)

cor(data$`Value (in M€)`, data$`Wage (in K€)`)
# Removing outliers regarding the transfer value from the dataset allows to decrease the variance
```

#### Plot of the Value vs Remaining Contract Years
```{r}
ggplot(data) +
  aes(x=`Wage (in K€)`, y=`Value (in M€)`, colour=Field_Position) +
  geom_point() +
  geom_smooth(method='lm') +
  scale_x_continuous(limits=c(0,300)) +
  theme_bw()
```

As expected, the palyers' market value is positively correlated to the players' wages. Looking at the different field positions we see that, for a given market value, Defensers get a better wage than Attackers. This reflects the law of offer and supply: there are more Attackers valued at 30 M€ than Defensers so the latter get better paid.

#### Plot
```{r}
ggplot(data) +
  aes(x=`Offensive Work Rate`, y=`Value (in M€)`, fill=`Offensive Work Rate`) +
  geom_boxplot() +
  scale_y_log10()
```


#### Plot
```{r}
ggplot(data) +
  aes(group=Overall, y=`Value (in M€)`, fill=Overall) +
  geom_boxplot(colour="Black", size=0.3) +
  scale_fill_gradient(low="White", high="Blue") +
  scale_x_continuous(limits=c(-0.2,0.4)) +
  facet_grid(Field_Position~.)
```

#### Adding teams and leagues to database:
```{r}
topclub<-data %>% group_by(Club) %>% summarize(count=n(), Val=sum(`Value (in M€)`),) %>% arrange(desc(Val)) #TopClub by nb of player and Value in M€

teams<-read.csv("teams.csv")
league<-read.csv("leagues.csv")
league$LeagueCode<-League$Code
teams2<-merge(teams, league, by="LeagueCode", all= F)
teams2$Club<-teams2$Name.x
test<-merge(data, teams2, by="Club", all= T)
```

***

# III-Regression models : predict Value with inputs

#### 1. Linear regression (full)
```{r}
datareg <- data %>% select(-ID, -Name, -Nationality, -Club)

#We do a 10-folds cross validation
set.seed(12345)

n=nrow(data)
MSE=rep(NA,20)
MAPE=rep(NA,20)
for (i in 1:20){
  train_index=sample(1:n,round(0.7*n))
  train=datareg[train_index,]
  test=datareg[-train_index,]
  linear_model <- lm(Value (in M€)~., data=train)
  pred_lm=predict(linear_model,newdata=test)
  MSE[i]=mean((test$`Value (in M€)`-pred_lm)^2)
  MAPE[i]=mean(abs((test$`Value (in M€)`-pred_lm)/(test$`Value (in M€)`)))
  
}

MSE_lm=mean(MSE)
MAPE_lm=mean(MAPE)

print(MSE_lm);print(MAPE_lm)
```

#### 2. Trees and Random forests

```{r}
n<-nrow(data)
set.seed(12345)
train_index=sample(1:n,round(0.7*n))
train=datareg[train_index,]
test=datareg[-train_index,]

tree<-rpart("Value (in M€)"~., data=train, cp=0.000000001)
printcp(tree) 
plotcp(tree)
alpha.opt<-tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"]
besttree<-prune(tree, cp=alpha.opt)
pred<-predict(besttree, newdata=test)
mean((data$"Value (in M€)"-pred)^2) #quadratic err

visTree(besttree)
```
